An Experimental Analysis of the Influence of the Precision of Parameters in an Artificial Neural Network

### 2025-11-12 paper/main.tex
weiter

\begin{abstract}
	\noindent
	We study how the numerical precision of weights, activations, and gradients in
	neural networks can be reduced without sacrificing too much of the performance.
	Using MNIST as a dataset, we benchmark uniform quantization (mapping real
	values to evenly spaced levels via a single step size), mixed-precision
	formats, and post-training decimal rounding. We evaluate precision settings
	(FP64/FP32/FP16) across all tensors during training, experiment with an
	inference-only quantization path on FP64-trained models, and test automatic
	mixed precision with gradient clipping. Metrics include test accuracy,
	training/inference time, and parameter storage. We compare end-to-end
	low-precision training, inference-only quantization, and mixed-precision with
	gradient clipping, and evaluate their impact on test accuracy and efficiency.
	The results on the used test setup show that (i) 16-bit end-to-end training
	does not significantly decrease test accuracy for the MNIST dataset, (ii)
	three-decimal-place weight rounding while keeping 32-bit precision incurs
	minimal accuracy loss. These findings verify limits for memory- and
	energy-constrained usecases like inference in mobile devices.
\end{abstract}

>>>>

\begin{abstract}
	\noindent
	We study how the numerical precision of weights, activations, and gradients in
	neural networks can be reduced without sacrificing too much of the performance.
	Using MNIST as a general dataset (and FashionMNIST for the decimal rounding),
	we benchmark uniform quantization (mapping real values to evenly spaced levels
	via a single step size), mixed-precision formats, and post-training decimal
	rounding. We evaluate precision settings (FP64/FP32/FP16) across all tensors
	during training, experiment with an inference-only quantization path on
	FP64-trained models, and test automatic mixed precision with gradient clipping.
	Metrics include test accuracy, training/inference time, and parameter storage.
	We compare end-to-end low-precision training, inference-only quantization, and
	mixed-precision with gradient clipping, and evaluate their impact on test
	accuracy and efficiency. The results on the used test setup show that (i)
	16-bit end-to-end training does not significantly decrease test accuracy for
	the MNIST dataset, (ii) three-decimal-place weight rounding while keeping
	32-bit precision incurs minimal accuracy loss. These findings verify limits for
	memory- and energy-constrained usecases like inference in mobile devices.
\end{abstract}
